---
title: "5/28/2020"
author: "Anna Madison Burns"
date: "5/28/2020"
output: html_document
---

```{r setup, include=FALSE}
require(mosaic)
require(lubridate)
require(MASS)
```

```{r}
NO2_R <- NO2_CA[ which(NO2_CA$`County Name`=='Riverside'), ]
```

```{r}
NO2_R <- NO2_R[c(-1,-2, -3, -4, -5, -6, -7, -8, -12, -13, -15, -16, -17, -18, -19, -20, -21, -22, -23, -24)]
```

```{r}
colnames(NO2_R) <- c("type","date","time","sample")
```

```{r}
NO2_R$type[NO2_R$type=="Nitrogen dioxide (NO2)"] <- "NO2"
```

```{r}
NO2_R$light <- "Day"
NO2_R$light[NO2_R$time<hms("08:00:00")|NO2_R$time>hms("19:00:00")] <- "Night"
```

```{r}
ggplot(data=NO2_R, mapping=aes(time, sample))+geom_point()
```

```{r}
NO2_R$traffic <- "Minimal"
NO2_R$traffic[NO2_R$time>hms("06:59:00") & NO2_R$time<hms("09:01:00")|NO2_R$time>hms("15:59:00")&NO2_R$time<hms("18:01:00")] <- "Heavy"
```

```{r}
fit1 <- lm(sample~light+time+date+traffic, data=NO2_R)
summary(fit1)
plot(fit1)
```

Standardized residuals are not great... need to transform.  Sqrt is much better... would like to test out log/look at a boxcox but it won't let me.  Why?  (Says response variables need to be positive... do I have any zero measurements?)

-> issues NA.rm=true, OR add very very small number to everything

Transforming data to make everything positive (adjusting for errors in data -> can't be a negative amount of NO2 measured! but left in as part of QAQC)

```{r}
NO2_R$sample <- (NO2_R$sample+0.15)
```

```{r}
boxcox(fit1)
```

Log appears to be an appropriate transformation for "sample".

```{r}
fit2 <- lm(log(sample)~light+time+date+traffic, data=NO2_R)
summary(fit2)
plot(fit2)
```

Way better! Awesome!

```{r}
NO2_R$sample <- log(NO2_R$sample)
```

Now beginning AIC...

```{r}
fit2 <- lm(sample~date, data=NO2_R)
summary(fit2)
```

This is clearly pretty good, and explains most of the correlation in our full model... keep it.  -> also indicates that we should look some more at daily variances... temperature, rain, etc.

```{r}
fit3 <- lm(sample~date+time, data=NO2_R)
summary(fit3)
anova(fit2,fit3)
```

Time explains variance that isn't explained by date, so both are significant - maybe specify "light" more -> peaks seem to be at 8am/5pm (heavy traffic!) -> specify traffic variable.  "Heavy traffic" from 7 am - 9 am, 4 am - 6 pm.

```{r}
fit4 <- lm(sample~traffic+time+date, data=NO2_R)
summary(fit4)
anova(fit3, fit4)
```

This is good! Do time and traffic levels interact?

```{r}
fit5 <- lm(sample~time+traffic+date+traffic*time, data=NO2_R)
anova(fit4,fit5)
```

Nope, no real interaction.  That makes sense, because we parsed "traffic" from "time".  

Let's look at one last thing related to time, and then move on to breaking down daily changes (temperature, rainfall, etc.)

```{r}
fit6 <- lm(sample~time+traffic+date+light, data=NO2_R)
summary(fit6)
anova(fit4,fit6)
```

Looks like we should keep it! And awesome news, we have half the variability explained.

Temperature, rainfall, cloudcover -> deeper level is wind, spatial modeling

Merging with data from the Weather Channel (weather) and UC Riverside (temperature, rainfall) ->



-> -> I need to somehow make each date's weather report repeat 24 times so that the rows line up with NO2_R, and it's able to bind.

```{r}
#nrow(Weather_R) == nrow(Weather_R) * 24
#Weather_R <- data.frame(a = 1:31, b = letters[1:31]) Weather_R[rep(seq_len(nrow(Weather_R)), each = 24), ]
#nrow(NO2_R) == nrow(Weather_R) * 24
#rep(Weather_R, 24)
Weather_R <- rep(Weather_R, each=24)
```

```{r}
NO2_R2 <- rbind(NO2_R, Weather_R, date)
```

