---
title: "Riverside Air Quality Analysis"
author: "Anna Madison Burns"
date: "5/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
require(mosaic)
require(lubridate)
```
##Updates:
AB 5/28 15:46 - Fixed it so that it would knit, and organized the data more nicely.  Keeping this markdown because it may be useful at some point to have the composite data, but am doing current work in the individual NO2/PM10 files so that I can create linear regressions.

# Work plan:
- establish hypothesis
- subset data (whole dataset is too large)
- visualize subset (need still smaller - a few plots of time vs. pollutant level)
- specify model to test hypothesis (pollution is a function of time of the day -> day vs. night lots of pollutants are UV specific -> figure out how long the sun is out by day, merge a different dataset/create categorical variable that assigns 8 pm - 7am of night, but can have more potential for analyzing data)
- run model and diagnostics (look at assumptions, compare models w/ AIC)

# Introduction

## Background

## Objectives

# Methods

## Data Processing

```{r, include=FALSE}
library(readr)
NO2_CA <- read_csv("~/Summer Research 2020/NO2_CA.csv")
View(NO2_CA)

library(readr)
PM10_CA <- read_csv("~/Summer Research 2020/PM10_CA.csv")
View(PM10_CA)
```

Source of data?

need to read the data for the Rmd file to work...

## Subsetting the data

```{r}
NO2_R <- NO2_CA[ which(NO2_CA$`County Name`=='Riverside'), ]
```

```{r}
PM10_R <- PM10_CA[ which(PM10_CA$`County Name` == 'Riverside'),]
```

```{r}
Riverside <- rbind(PM10_R, NO2_R)
```

```{r}
Riverside <- Riverside[c(-1,-2, -3, -4, -5, -6, -7, -8, -12, -13, -15, -16, -17, -18, -19, -20, -21, -22, -23, -24)]
```

```{r}
colnames(Riverside) <- c("type","date","time","sample")
```


```{r}
Riverside$light <- "Day"
  #format(Riverside$`Time Local`, format="%H%M%S")
Riverside$light[Riverside$time<hms("08:00:00")|Riverside$time >hms("19:00:00")] <- "Night"
  # Riverside$`Light` = with(Riverside, if(`Time Local`< "08:00:00" & `Time Local` > "20:00:00")"Night")
# str(Riverside)
```

Other Options:

flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )


x$daynight = with(x, ifelse(TOD > "053000" & TOD < "200000", "Day", "Night")) 

OR Light Function

Specify Model:

```{r}
Riverside$type[Riverside$type=="Nitrogen dioxide (NO2)"] <- "NO2"
Riverside$type[Riverside$type=="PM10 - LC"] <- "PM10LC"
Riverside$type[Riverside$type=="PM10 Total 0-10um STP"] <- "PM10T"
```


```{r}
interaction.plot(Riverside$light, Riverside$type, Riverside$sample)
```

```{r}
ggplot(data=Riverside, mapping=aes(time, sample)) + geom_point(mapping=aes(color=type))
```
